---
output: github_document
editor_options: 
  chunk_output_type: console
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# tweedekamer

<!-- badges: start -->
<!-- badges: end -->

This package provides bindings for using the [Open Data API of the Dutch House of Representatives](https://opendata.tweedekamer.nl/documentatie/odata-api)

## basic example code

## Installation

You can install the development version of tweedekamer from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("kasperwelbers/tweedekamer")
```

## Basic Usage

List entities that can be queried

```{r}
library(tweedekamer)

tk_entities()
```
Get attributes and relations for an entity

```{r}
tk_attributes("persoon")
```

Count items for an entity

```{r example}
tk_entity('persoon') |>
  tk_count()
```

Filter results

```{r}
tk_entity('persoon') |>
  tk_filter(
    by_string('geslacht', 'eq', 'vrouw')
  ) |>
  tk_count()
```

Download results

```{r}
data = tk_entity('persoon') |>
  tk_filter(
    by_string('geslacht', 'eq', 'vrouw')
  ) |>
  tk_list()

nrow(data)
```

Fancy filters

```{r}
tk_entity('persoon') |>
  tk_filter(
    by_string('geslacht', 'eq', 'vrouw'),
    filter_or(
      by_date('geboortedatum', 'lt', '1970-01-01T23:00:00Z'),
      by_date('geboortedatum', 'gt', '1980-01-01')
    )
  ) |>
  tk_count()
```

Expand (join) entities

```{r}
tk_entity('ZaakActor') |>
  tk_filter(by_string('ActorFractie', 'eq', 'VVD'), 
            by_string('relatie', 'eq', 'Indiener')) |>
  tk_expand('Zaak') |>
  tk_list(limit=5)
```




